Genetic Dilution erodes productivity: Exploring farmers' low adoption levels of improved maize in Ethiopia

### 2023-03-30 paper.tex
Version submitted to JDE

    \title{Genetic Dilution erodes productivity: Exploring farmers' low adoption levels of improved maize in  Ethiopia}
% This second title is my preferred title at this point (28 Oct 2021). Open to other titles, but it shouldn't be longer than this one!

@@

    \clearpage\maketitle
	\thispagestyle{empty}
	\vspace*{-2em}
    \begin{center}\begin{abstract}
    \singlespacing
			\noindent  
			Despite large investments in developing new maize varieties, evidence shows that few farmers consistently adopt them. Average effects on yields mask considerable heterogeneity. Misclassification of seed is a crucial problem in these contexts, as non-classical measurement error can lead to biased effects. We study misclassification in the context of heterogeneous impacts to adoption. We use a three-year panel to estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find results driven by bias due to the seed misclassification. We leverage unique information on maize DNA-fingerprinting collected over the same areas in 2018. Misclassification in self-reported hybrid seed adoption masks heterogeneity in the quality of the genetic material categorized as improved varieties. Accounting for seed misclassification, we find that heterogeneity in the actual quality of the genetic material masks positive returns to adoption, with the highest returns found for farmers using higher-purity germplasm, drought-tolerant maize, and newly released varieties. When we account for misclassification in our heterogeneity analysis, we find that farmers that have the largest potential to benefit from adoption are constrained in their ability to do so. The wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium points to the need for policies that expand the accessibility of improved seed and verify its quality.

@@

        {\small \noindent\emph{JEL Classification}: O12, O13, O33, Q12, C13 \\
        	\emph{Keywords}: Improved maize, DNA-fingerprinting, Group Random Coefficient, Endogenous Switching Regression, Heterogeneous returns}

>>>>

    \title{Genetic Dilution erodes productivity: Exploring farmers' low adoption levels of improved maize in  Ethiopia\footnote{We are grateful for comments on earlier drafts of the paper from Travis Lybbert and James Stevenson, as well as comments from participants in presentations at NEUDC 2022. This research was supported by the CGIAR Standing Panel on Impact Assessment (SPIA) under its program of work 2019-2021. Thank you to Mai Mahmoud, Mina Karasalo, Karen Macours and the rest of the SPIA Small Grants team. All errors are our own. Declarations of interest: none.} }
% This second title is my preferred title at this point (28 Oct 2021). Open to other titles, but it shouldn't be longer than this one!

@@

    \clearpage\maketitle
	\thispagestyle{empty}
	\vspace*{-2em}
    \begin{center}\begin{abstract}
    \singlespacing
			\noindent
   Household surveys often rely on self-reports of hybrid seed use, which are prone to measurement error. This context exhibits considerable heterogeneity when estimating productivity impacts. Using a three-year panel, we estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find results driven by non-classical measurement error due to seed misclassification. We leverage information on DNA-fingerprinting collected over the same areas in 2018 and find that heterogeneity in the quality of the genetic material masks positive returns to adoption, with the highest returns found for farmers using higher-purity, drought-tolerant, and newly released varieties. When we account for misclassification in our initial heterogeneity analysis, we find that farmers that have the largest potential to benefit from adoption are constrained in their ability to do so. Our research points to the need for policies that expand the accessibility of improved seed and the need for mechanisms to verify its quality.
			% Despite large investments in developing new maize varieties, evidence shows that few farmers consistently adopt them. Average effects on yields mask considerable heterogeneity. Misclassification of seed is a crucial problem in these contexts, as non-classical measurement error can lead to biased effects. We study misclassification in the context of heterogeneous impacts to adoption. We use a three-year panel to estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find results driven by bias due to the seed misclassification. We leverage unique information on maize DNA-fingerprinting collected over the same areas in 2018. Misclassification in self-reported hybrid seed adoption masks heterogeneity in the quality of the genetic material categorized as improved varieties. Accounting for seed misclassification, we find that heterogeneity in the actual quality of the genetic material masks positive returns to adoption, with the highest returns found for farmers using higher-purity germplasm, drought-tolerant maize, and newly released varieties. When we account for misclassification in our heterogeneity analysis, we find that farmers that have the largest potential to benefit from adoption are constrained in their ability to do so. The wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium points to the need for policies that expand the accessibility of improved seed and verify its quality.

@@

        {\small \noindent\emph{JEL Classification}: O12, O13, O33, Q12, C13 \\
        	\emph{Keywords}: Improved maize, DNA-fingerprinting, Group Random Coefficient, Endogenous Switching Regression, Heterogeneous returns, Ethiopia}
